---
title: "Homework2_201611531"
author: "Jeong Hojae"
date: '2021 3 27 '
output:
  pdf_document: default
  html_document: default
---

## 0. Read the data

```{r data}
library(readr)
library(ggplot2)
library(forecast)
library(tseries)

spot <- read_csv("/cloud/project/spot.csv")
dim(spot)
summary(spot)
boxplot(spot)
plot(c(1:100), t(spot),type = 'l')
```

## 1. Draw time series graph of sun spot. Do you see any cycle or seasonal effect?

```{r num_1}
spot.ts=ts(data=spot, frequency=1)
plot(spot.ts, mai ="Time Series graph of Sun spot")
abline(h=mean(spot.ts[,1]),col="blue")

tseries::kpss.test(spot.ts,null="Level")
tseries::kpss.test(spot.ts,null="Trend")
```

H0 : Data is level stationary. (tseries::kpss.test(spot.ts, null="Level"))

When significance level is 0.05, p-value (0.1) is greater than it.
So, We cannot reject H0.


H0 : Data is trend stationary. (tseries::kpss.test(spot.ts, null="Trend"))

When significance level is 0.05, p-value (0.0402) is smaller than it.
So, We can reject H0.

Therefore, This data has level.
Alse, we can see cycle or seasonal effect easily at the graph.


## 2. try 5-point moving average smoothing. Draw the plot of original graph in black, 5 point MA smoothing in red, mean value in blue.

```{r num_2}
m5=filter(spot.ts, filter=rep(1/5,5), method="convolution", sides=1)
plot(spot.ts,main="5-point moving average")
lines(m5,col="red",lty=1, lwd=2)
abline(h=mean(spot.ts[,1]),col="blue")
```

There is no center to calculate average.

$M_5=[Y_5+Y_4+Y_3+Y_2+Y_1]/5$


## 3. Check the residual plot, check the stationary and the test the independence assumption. Carefully interpret the residual analysis.

```{r num_3_1}
head(m5,10)
res=m5[-1:-4,]-spot.ts[-1:-4,]
head(res,10)
tsdisplay(res, main="residuals by MA(5) smoothing")
```

At the residual plot, it seems to have pattern.

Also, several bars are crossing the blue line (significance level) at the ACF and PACF graphs.

Therefore, The residual shows auto correlation after the model MA(5) is applied. i.e. It is non-stationary.

```{r num_3_2}
Box.test(res)
```
H0 : given time series data is independence

When significance level is 0.05, p-value (5.044e-13) is smaller than it.
So, We can reject H0.

Therefore, independence of residuals doesn't exist.


## 4. Fit the simple exponential smoothing with alpha=0.1 and with the optimized alpha. If you think we need a trend, or seasonal, or both try them. Please address all the modeling and show how you find the best exponential smoothing model for spot data.

Let's compare exponential smoothing in all cases.

### no trend and no seasonal effect

```{r num_4_a_1}
ho=HoltWinters(spot.ts, alpha=0.1, beta=F, gamma=F) 
#exponential smoothing # (beta=F, gamma=F): no trend and no seasonal effect
ho
plot(ho)

ha=HoltWinters(spot.ts,beta=F, gamma=F) #exponential smoothing
ha
plot(ha) # the red line is the fitted value
```
```{r num_4_a_2}
fo=forecast(ho)
tsdisplay(fo$residual)
Box.test(fo$residual, type="Box-Pierce")

fa=forecast(ha)
tsdisplay(fa$residual)
Box.test(fa$residual, type="Box-Pierce")
```
```{r num_4_a_3}
accuracy(fo)
accuracy(fa)
```
### no trend effect

To consider seasonal variation, I set up frequency to 4.

```{r num_4_b_1}
spot.ts4=ts(data=spot,frequency=4) # seasonal data
ho=HoltWinters(spot.ts4, alpha=0.1, beta=F) #exponential smoothing # (beta=F): no trend effect
ho
plot(ho)

ha=HoltWinters(spot.ts4,beta=F) #exponential smoothing
ha
plot(ha) # the red line is the fitted value
```

```{r num_4_b_2}
fo=forecast(ho)
tsdisplay(fo$residual)
Box.test(fo$residual, type="Box-Pierce")

fa=forecast(ha)
tsdisplay(fa$residual)
Box.test(fa$residual, type="Box-Pierce")
```

```{r num_4_b_3}
accuracy(fo)
accuracy(fa)
```

### no seasonal effect

```{r num_4_c_1}
ho=HoltWinters(spot.ts, alpha=0.1, gamma=F) #exponential smoothing # (gamma=F): no seasonal effect
ho
plot(ho)

ha=HoltWinters(spot.ts, gamma=F) #exponential smoothing
ha
plot(ha) # the red line is the fitted value
```

```{r num_4_c_2}
fo=forecast(ho)
tsdisplay(fo$residual)
Box.test(fo$residual, type="Box-Pierce")

fa=forecast(ha)
tsdisplay(fa$residual)
Box.test(fa$residual, type="Box-Pierce")
```

```{r num_4_c_3}
accuracy(fo)
accuracy(fa)
```

### trend and seasonal effect exist

```{r num_4_d_1}
ho=HoltWinters(spot.ts4, alpha=0.1) #exponential smoothing
ho
plot(ho)

ha=HoltWinters(spot.ts4) #exponential smoothing
ha
plot(ha) # the red line is the fitted value
```

```{r num_4_d_2}
fo=forecast(ho)
tsdisplay(fo$residual)
Box.test(fo$residual, type="Box-Pierce")

fa=forecast(ha)
tsdisplay(fa$residual)
Box.test(fa$residual, type="Box-Pierce")
```

```{r num_4_d_3}
accuracy(fo)
accuracy(fa)
```

RMSE of the optimized alpha is always smaller than RMSE of alpha=0.1. 

In case of only no trend effect and optimized alpha, residual plot doesn't seem to have pattern.

In all cases, several bars are crossing the blue line at ACF and PACF graph.

Plus, p-value is smaller than significance level (0.05) at Box-Pierce test.

Therefore, residuals isn't stationary and independence.


To find the best exponential smoothing model for spot data, We must compare RMSE of all models.
RMSE was the lowest in the model with no trend effect, no seasonal effect and optimized alpha (0.9999339).

## 5.From your best model, find the forecast of next 4 points.

```{r num_5}
ha=HoltWinters(spot.ts,beta=F, gamma=F) #exponential smoothing
fa=forecast(ha, h=4)
fa
plot(fa,main="80%, 95% significant level for forecasting") 
lines(fa$fitted, col="red", lty=2, lwd=2)


Fa=forecast(ha, h=4, fan=T)
Fa
plot(Fa,main="51-99% significant level for forecasting") 
lines(Fa$fitted,col="red", lty=2, lwd=2)
```

The next 4 points are all 73.99755 from my best model.