---
title: "Homework3_201611531"
author: "Jeong Hojae"
date: '2021 4 4 '
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## 0. Read the data

```{r data}
library(tidyverse)
library(ggplot2)
library(forecast)
library(seasonal)
library(aTSA)
library(timsac)

data <- read_csv("/cloud/project/loadregr.csv")
dim(data)
summary(data)
head(data)
data_ts<-ts(data$MKw, start=c(1970, 5), end=c(1979, 12), frequency=12 )
```

## 1. Draw time series plot

```{r 1}
data_ts %>% autoplot() +
  ggtitle("Time Series Graph of The Monthly Peak Load for Electricity Iowa")

data_ts %>%
  stl(t.window=13, s.window="periodic", robust=TRUE) %>%
  autoplot()+
  ggtitle("Some graphs")
```

The data has trend variation and seasonal trend. But, it doesn't seem to show about dependent of them.

## 2. Fit decomposition model and draw a plot with the original data, seasonally adjusted, trend-cycle component all together.

### Classical decomposition
```{r Classical_additive}
# additive
dd1=decompose(data_ts,type="additive")
tseries::kpss.test(dd1$random,null="Level")
ggtsdisplay(dd1$random, main="residual")

autoplot(data_ts, series="Data")+
  autolayer(trendcycle(dd1),series="Trend") +
  autolayer(seasadj(dd1),series="Seasonally Adjusted") +
  xlab("Time") + ylab("New orders index")+
  ggtitle("Classical decomposition(additive)") +
  scale_color_manual(values=c("gray","blue","red"),breaks=c("Data","Seasonally Adjusted","Trend"))

```

Additive model is used when the seasons are constant regardless of the trend


```{r Classical_multiplicative}
# multiplicative
dd2=decompose(data_ts,type="multiplicative")
tseries::kpss.test(dd2$random,null="Level")
ggtsdisplay(dd2$random, main="residual")

autoplot(data_ts, series="Data")+
  autolayer(trendcycle(dd2),series="Trend") +
  autolayer(seasadj(dd2),series="Seasonally Adjusted") +
  xlab("Time") + ylab("New orders index")+
  ggtitle("Classical decomposition(multiplicative)") +
  scale_color_manual(values=c("gray","blue","red"),breaks=c("Data","Seasonally Adjusted","Trend"))

```

Multiplicative model is used when the seasons change with the trend.


In classical decomposition, the first and last few observations of the trend-cycle factor is removed.

So, Let's apply another method. The following three methods don't make data loss.

Plus, We cannot reject H0 in kpss.test. And the value is usually with in the significance level in the ACF and PACF graph. So, Data is stationary.

### SEATS decomposition
```{r SEATS}
dd3<-seas(data_ts)
names(dd3)
dd3_r<-remainder(dd3)
tseries::kpss.test(dd3_r,null="Level")
ggtsdisplay(dd3_r, main="remainder")

autoplot(data_ts, series="Data")+
  autolayer(trendcycle(dd3),series="Trend") +
  autolayer(seasadj(dd3),series="Seasonally Adjusted") +
  xlab("Time") + ylab("New orders index")+
  ggtitle("SEATS decomposition") +
  scale_color_manual(values=c("gray","blue","red"),breaks=c("Data","Seasonally Adjusted","Trend"))
```

The procedure works only with quarterly and monthly data.

There is no problem because the data is monthly.

Plus, We cannot reject H0 in kpss.test. And the value is usually with in the significance level in the ACF and PACF graph. So, Data is stationary.

### X11 decomposition
```{r 2-4}
data_ts %>% seas(x11="") ->dd4
dd4_r=remainder(dd4)
tseries::kpss.test(dd4_r,null="Level")
ggtsdisplay(dd4_r, main="remainder")

autoplot(data_ts, series="Data")+
  autolayer(trendcycle(dd4),series="Trend") +
  autolayer(seasadj(dd4),series="Seasonally Adjusted") +
  xlab("Time") + ylab("New orders index")+
  ggtitle("X11 decomposition") +
  scale_color_manual(values=c("gray","blue","red"),breaks=c("Data","Seasonally Adjusted","Trend"))
```

Trend-cycle estimates are also available for all observations including the end points.

And the seasonal component is allowed to vary slowly over time.

So, that line isn't flatter than other procedure.

Plus, We cannot reject H0 in kpss.test. And the value is usually with in the significance level in the ACF and PACF graph. So, Data is stationary.

### STL decomposition
```{r 2-5}
dd5<-stl(data_ts,'periodic')
dd5_r=remainder(dd5)
tseries::kpss.test(dd5_r,null="Level")
ggtsdisplay(dd5_r, main="remainder")

autoplot(data_ts, series="Data")+
  autolayer(trendcycle(dd5),series="Trend") +
  autolayer(seasadj(dd5),series="Seasonally Adjusted") +
  xlab("Time") + ylab("New orders index")+
  ggtitle("STL decomposition") +
  scale_color_manual(values=c("gray","blue","red"),breaks=c("Data","Seasonally Adjusted","Trend"))
```

STL will handle any type of seasonality, not only monthly and quarterly data.

But, it does not handle trading day or calendar variation automatically, and it only provides facilities for additive decompositions.

Plus, We cannot reject H0 in kpss.test. And the value is usually with in the significance level in the ACF and PACF graph. So, Data is stationary.

Therefore, X11 seems most appropriate method considering the shape and stationary of the data.

## 3. Draw sub-series seasonal effect graph in each month.

```{r 3}
dd4 %>% seasonal() %>% ggsubseriesplot()+ylab("Seasonal")
```

This graph shows shape of seasonal effect of each month
(Blue line is constant seasonal effect of each month)

The monthly peak load for electricity Iowa is high around May and November.
And it is low around the end of January and the end of January

## 4. Draw a polar seasonal plot.

```{r 4}
ggseasonplot(data_ts, polar=TRUE) +
  ylab("MKw") +
  ggtitle("Polar seasonal plot")
```

The graph show not only just the same results at the graph of number three but also an increase over the years.



