---
title: "Homework4_201611531"
author: "Jeong Hojae"
date: '2021 4 20 '
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
library(tidyverse)
library(ggplot2)
library(forecast)
library(gridExtra)
```

## 1. Simulate an AR(1) process with phi=0.8 (n=1000) and show the time series plot, SACF graph and SPACF graph.

```{r sol1}
y1=arima.sim(n=1000, list(ar=0.8))
autoplot(y1) + ggtitle("AR(1) with phi=0.8 ")
ggAcf(y1)+ggtitle("SACF with phi=0.8")
ggPacf(y1)+ggtitle("SPACF with phi=0.8")
```

Time series plot don't get out of average. And it shows high relation with previous data value. It is stationary time series.

SACF graph shows exponential decline.

SPACF graph cuts-off after 2nd lag.



## 2. Repeat (1) process in case of phi=-0.8.

```{r sol2}
y2=arima.sim(n=1000, list(ar=-0.8))
autoplot(y2) + ggtitle("AR(1) with phi=-0.8 ")
ggAcf(y2)+ggtitle("SACF with phi=-0.8")
ggPacf(y2)+ggtitle("SPACF with phi=-0.8")
```

Time series plot don't get out of average. And it shows high relation with previous data value. It is stationary time series. Also, there are more to vibrate than 1 graph because phi is negative number.

SACF graph vibrates and decreases exponential decline.

SPACF graph vibrates and cuts-off after 2nd lag.


## 3. Simulate an AR(2) process with randomly chosen phi1 and phi2. And Repeat (1) process.


```{r sol3}
set.seed(201611531)
phi1=runif(1,-1,1)
phi2=runif(1,-1,1)
phi1
phi2


y3<- arima.sim(n=1000, list(ar=c(phi1, phi2)))
autoplot(y3) + ggtitle("AR(2) with phi1=0.172035, phi2=0.5103432")
ggAcf(y3)+ggtitle("SACF with phi1=0.172035, phi2=0.5103432")
ggPacf(y3)+ggtitle("SPACF with phi1=0.172035, phi2=0.5103432")
```

Time series plot don't get out of average. And it shows high relation with previous data value. It is stationary time series. Also, there are more to vibrate than 1 graph but it's not more to vibrate than 2 graph.

SACF graph decreases exponential decline. It has a real root.

SPACF graph cuts-off after 3rd lag because it's AR(2) model.




