---
title: "time series-Module9"
author: "Jeong Hojae"
date: '2021 5 20 '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Load libraries

```{r library, warning=F, message=F}
library(tidyverse)
library(ggplot2)
library(forecast)
library(gridExtra)
library(tseries)
library(fUnitRoots)
```


## Augmented Dickey-Fuller tests

The Augmented Dickey-Fuller test allows for higher-order autoregressive processes by including ∆yt−p in the model. But our test is still if $\gamma = 0.$
$$\Delta y_t = α + βt + γy_{t−1} + δ_1∆y_{t−1} + δ_2∆y_{t−2} + . . .$$
The null hypothesis for both tests is that the data are non-stationary (the unitroot exists). We want to REJECT the null hypothesis for this test, so we want a p-value of less that 0.05 (or smaller).


##ADF test using `adf.test()` from `tseries`.

```{r adf-wn}
TT <- 100
wn <- rnorm(TT) # white noise
autoplot(ts(wn))
tseries::adf.test(wn)
```

The null hypothesis is rejected.

단위근이 있다는 귀무가설이 기각된다.

Try a Dickey-Fuller test. This is testing with a null hypothesis of AR(1) stationarity versus a null hypothesis with AR(4) stationarity when we used the default k.

따라서 단위근이 없는 계열로 확인되었다.





More restrictive model testing.

```{r adf-wn2}
tseries::adf.test(wn, k = 0) # order=0
```

Notice that the test-statistic is smaller. This is a more restrictive test and we can reject the null with a higher significance level.



## Test Test on white noise with trend
Try the test on white noise with a **trend** and **intercept**.


```{r adf-wnt}
intercept <- 1
wnt <- wn + 1:TT + intercept
autoplot(ts(wnt))
tseries::adf.test(wnt)
#?adf.test

```

The null hypothesis is still rejected. adf.test() uses a model that allows an intercept and trend. (귀무가설 기각 단위근이 없다.)

시간에 의해서 증가하는 그래프 trend 존재


## Test with a random walk

```{r adf-rw}
random.walk <- ts(cumsum(wn))
autoplot(random.walk)
adf.test(random.walk)


```

This p-value is high, so we fail to reject the null hypothesis. This is a non-stationary series.


## The monthly peak load for electricity Iowa

Please address following four questions using the data of the electricity data at Iowa state.

```{r adf1}
data<-read.csv("/cloud/project/loadregr.csv") #find a proper directory for your data.
data_ts<-ts(data$MKw, start=c(1970, 5), end=c(1979, 12), frequency=12 )
autoplot(data_ts)
ggAcf(data_ts)
ggPacf(data_ts)


tseries::adf.test(data_ts)
```

we reject the null hypothesis. This is a stationary series.

```{r adf2}
aic=c()


#p=1
for(p in 1:10){
  ar.fit=Arima(data_ts, order=c(p,0,0))
  aic[p]=ar.fit$aic
}


# save ar.fit
ar.fit<-list()

for(p in 1:10){
  ar.fit[[p]]=Arima(data_ts, order=c(p,0,0))
  aic[p]=ar.fit[[p]]$aic
}

ar.fit[[3]]

#names(ar.fit)
#ar.fit$sigma2


P<-1:10
aic_data<-data.frame(P, aic)
aic_data

ggplot(data=aic_data, aes(x=P, y=aic)) +
  geom_line(linetype = "dashed")+
  geom_point()

which.min(aic) #9


fUnitRoots::adfTest(data_ts, type="ct", lags=8) #ct: intercept with time trend
# lag=aic_order-1
tseries::adf.test(data_ts, k=8)
tseries::adf.test(data_ts)


```

- fUnitRoots::adfTest(data_ts, type="ct", lags=8) 

p-value 작으므로 귀무가설을 기각, 단위근이 없는 계열

- tseries::adf.test(data_ts, k=8)

마찬가지의 결과가 나옴

- tseries::adf.test(data_ts)

k의 defalt 4를 사용, 이 때도 귀무가설을 기각하므로 정상시계열



## Airpassenger data
Let’s do one more example — this time the well-known monthly airpass series, which has both trend and seasonality. So this won’t be stationary.

```{r adf3}

library(aTSA)
autoplot(AirPassengers)
ggAcf(AirPassengers)
ggPacf(AirPassengers)
```


We know from the previous example that we need to set the lag order large enough for the test to see the seasonality. This is **monthly data** so we’ll choose k=12:

```{r adf4}
tseries::adf.test(AirPassengers, k=12)
```

As expected, we fail to reject the null hypothesis. This is a non-stationary series.

귀무가설을 기각하지 못하므로 단위근이 있는 모형이다.


## Conclusion

In this post we learned how to use the Augmented Dickey-Fuller test to determine whether a given time series is stationary. We did this using the adf.test() function in tseries package and adfTest() function in fUnitRoots package

