---
title: "timeserise_1"
author: "hojae"
date: '2021 3 8 '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Module-1 Time series practice

There are five time series data set. The first data set (dd1) is a random variation time series

```{r dd1}
#dd1

dd1=matrix(c(1342,1442,1252,1343,1425,1362,1456,1272,1243,1359,1412,1253,1201,1478,1322,1406,1254,1289,1497,1208))
dd1.ts=ts(data = dd1, start = c(2006,1), frequency = 4)
dd1.ts
plot(dd1.ts, mai='Random variation Time Series')

```


The second data set (dd2) is a seasonal variation time series
```{r dd2}
#dd2

dd2=matrix(c(1142,1242,1452,1543,1125,1262,1456,1572,1143,1259,1462,1553,1121,1258,1472,1546,1154,1249,1477,1548))
dd2.ts=ts(data = dd2, start = c(2006,1), frequency = 4)
dd2.ts
plot(dd2.ts, mai='Seasonal variation Time Series')
```


The third data set (dd3) is a trend variation time series
```{r dd3}
#dd3

dd3=matrix(c(1142,1242,1252,1343,1225,1562,1356,1572,1343,1459,1412,1453,1401,1478,1322,1606,1554,1589,1597,1408))
dd3.ts=ts(data = dd3, start = c(2006,1), frequency = 4)
dd3.ts
plot(dd3.ts, mai='Trend variation Time Series')
```


The fourth data set (dd4) is a seasonal and trend variation time series
```{r dd4}
#dd4

dd4=matrix(c(1142,1242,1452,1543,1225,1362,1556,1672,1343,1459,1662,1753,1421,1558,1722,1846,1554,1649,1877,1948))
dd4.ts=ts(data = dd4, start = c(2006,1), frequency = 4)
dd4.ts
plot(dd4.ts, mai='Seasonal Trend variation Time Series')
```


The last data set (dd5) is a cyclinical variation time series
```{r dd5}
#dd5

dd5=matrix(c(1142,1242,1452,1543,1225,1362,1556,1672,1343,1459,1662,1753,1221,1358,1572,1646,1154,1249,1477,1548))
dd5.ts=ts(data = dd5, start = c(2006,1), frequency = 4)
dd5.ts
plot(dd5.ts, mai='Cyclinical variation Time Series')
```

## Stationary

Let's generate a white nose series.

```{r WN}
WN <- runif(100, min = -0.5, max = 0.5)
WN.ts <- ts(data=WN, start = c(1,1), frequency = 1)
plot(WN.ts, mai="White noise")
abline(h=0)
```


The ACF and PACF plot show that the generated WN.ts is stationary.
```{r stationary}
acf(WN.ts, main="ACF")
pacf(WN.ts, main="ACF")
```


Testing if wn.ts is satisfying "stationary"
```{r stationary-test}
tseries::kpss.test(WN.ts, null="Level")
```
H0 : Data has Stationarity when except Level
p-value = 0.1
=> We cannot reject H0



## Non-Stationary

Let's generate a random walk series, which is non-stationary

```{r RW}
at <- runif(100, min = -0.5, max = 0.5)
yt <- runif(100, min = -0.5, max = 0.5)
for(i in 2:100){
  yt[i]=yt[i-1]+at[i]
}

yt.ts <- ts(data=yt)
plot(yt.ts, mai="Random walk")
abline(h=0)
```


The ACF and PACF plot show that the generated random walk process (yt.ts) is non stationary.
```{r non-stationary}
acf(yt.ts, main="ACF")
pacf(yt.ts, main="ACF")
```


KPSS test for stationary
```{r nonstationary-test}
tseries::kpss.test(yt.ts, null="Level")
tseries::kpss.test(yt.ts, null="Trend")
```
H0 : Data has level stationary. (tseries::kpss.test(yt.ts, null="Level"))
p-value = 0.1
=> We cannot reject H0

H0 : Data has trend stationary. (tseries::kpss.test(yt.ts, null="Trend"))
p-value = 0.01
=> We can reject H0, it has trend






