---
title: "time series3"
author: "hojae"
date: '2021 3 28 '
output: html_document
---

```{r tidyverse, warning=F, message=F}

# load libraries
#install.packages('seasonal')
#install.packages('fpp')
library(tidyverse)
library(ggplot2)
library(forecast)
library(fpp)
library(aTSA)
library(seasonal)
library(timsac)
```

```{r pipe line}
#pipe line (%>%) %>% Ctrl+Shift+M 

round(sqrt(85),2)

sqrt(85) %>% round(2)
```

In computing, a pipeline, also known as a data pipeline, is a set of data processing elements connected in series, where the output of one element is the input of the next one. The elements of a pipeline are often executed in parallel or in time-sliced fashion.


```{r data}

autoplot(AirPassengers) # easily draw TSA plot

AirPassengers %>% autoplot()

# plotting different way
AirPassengers %>%
  stl(t.window=13, s.window="periodic", robust=TRUE) %>%
  autoplot()

# without pipeline
?autoplot()
autoplot(stl(AirPassengers, t.window=13, s.window="periodic", robust=TRUE))
```

Pipeline can save the storage space.

# Classical decomposition
- The classical decomposition method originated in the 1920s.

- It is a relatively simple procedure, and forms the starting point for most other methods of time series decomposition.

- There are two forms of classical decomposition: an additive decomposition and a multiplicative decomposition.

- These are described below for a time series with seasonal period (e.g., $m = 4$ for quarterly data, $m = 12$
for monthly data, $m = 7$ for daily data with a weekly pattern).

- In classical decomposition, we assume that the seasonal component is constant from year to year.
For multiplicative seasonality, the values that form the seasonal component are sometimes called the “seasonal indices”.

## Decomposition of time series data - additive model
```{r decomp}


dd2=matrix(c(1142, 1242, 1452, 1543, 1125, 1262, 1456, 1572, 1143, 1259,
1462, 1553, 1121, 1258, 1472, 1546, 1154, 1249, 1477, 1548))
dd2.ts=ts(data=dd2, start=c(2006,1), frequency=4)


dd4=matrix(c(1142, 1242, 1452, 1543, 1225, 1362, 1556, 1672, 1343, 1459,
1662, 1753, 1421, 1558, 1772, 1846, 1554, 1649, 1877, 1948))
dd4.ts=ts(data=dd4, start=c(2006,1), frequency=4)

plot(decompose(dd2.ts, type="additive")) # additive model
plot(decompose(dd4.ts, type="additive")) # additive model


#decomposition
dd=decompose(dd4.ts, type="additive")

#decomposition results
names(dd)


dd$seasonal # seasonal effects

dd$trend # trend effects

dd$random #random variation effects

dd_noseasonal=dd4-dd$seasonal # remove seasonal effects

dd_notrend=dd4-dd$trend # remove trend effects


plot(dd_noseasonal, main="TS without seasonal effects")

plot(dd_notrend, main="TS without trend effects")

```


## Decomposition of time series data - multiplicative model

```{r decomp2}

#decomposition
dd_m=decompose(dd4.ts, type="multiplicative")

#decomposition results
names(dd_m)

dd_m$seasonal # seasonal effects

dd_m$trend # trend effects

dd_m$random #random variation effects

ddM_noseasonal=dd4/dd_m$seasonal # remove seasonal effects*

ddM_notrend=dd4/dd_m$trend # remove trend effects*


plot(ddM_noseasonal, main="TS without seasonal effects")

plot(ddM_notrend, main="TS without trend effects")

```


## Decomposition of Airpassenger data
Decompose Airpassenger data:


```{r airpass-decomp}
data("AirPassengers")
ddd<-decompose(AirPassengers, type="multiplicative")

plot(decompose(AirPassengers, type="additive"))
plot(decompose(AirPassengers, type="multiplicative"))
plot(AirPassengers/ddd$seasonal, main="Remove seasonal variation")
plot(AirPassengers/ddd$trend, main="Remove trend variation")


AirPassengers %>% decompose(type="additive") %>%
  autoplot() +
  ggtitle("Classical additive decomposition of airPassengers data")

AirPassengers %>% decompose(type="multiplicative") %>%
  autoplot() +
  ggtitle("Classical multiplicative decomposition of airPassengers data")

```

Diagnosis:

```{r airpass-diagnosis, warning=F, message=F}

dda=decompose(AirPassengers, type="additive")

ddm=decompose(AirPassengers, type="multiplicative")

tseries::kpss.test(dda$random, null = "Level") #Stationary for random variation after trend and seasonal effects were removed

tseries::kpss.test(ddm$random, null = "Level") #Stationary for random variation after trend and seasonal effects were removed

tsdisplay(dda$random, main="Random/ Additive model")
tsdisplay(ddm$random, main="Random/ Multiplicative model")
```

Additive and multiplicative decomposition model are stationary for random variation after trend and seasonal effects were removed. 

So, it's explaining trend and seasonal effects well.

When you test the stationary, you must check both `tseries::kpss.test` and `tsdisplay`.

`tseries::kpss.test` : stationary test
`tsdisplay` : ACF PACF graph


# SEATS decomposition
- “SEATS” stands for “Seasonal Extraction in ARIMA Time Series” (ARIMA models will be discussed
later).

- This procedure was developed at the Bank of Spain, and is now widely used by government agencies
around the world.

- The procedure works only with **quarterly and monthly** data. So seasonality of other kinds, such as daily data, or hourly data, or weekly data, require an alternative approach.

```{r SEATS, warning=F, message=F}
?seas()

data(elecequip)
autoplot(elecequip)

elecequip %>% seas() %>%
autoplot() +
  ggtitle("SEATS decomposition of electrical equipment index")

elecequip %>% seas()->seats_elece
```

# X11 decomposition

## X11 decomposition with r-packages with Tidyverse and ggplot2

- X11 decomposition is based on classical decomposition, but includes **many extra steps and features** in order to overcome the drawbacks of classical decomposition.

- In particular, trend-cycle estimates are available for all observations including the end points, and the seasonal component is allowed to vary slowly over time.

- X11 also has some sophisticated methods for handling trading day variation, holiday effects and the effects of known predictors.

- It handles both **additive and multiplicative decomposition**.

- The process is entirely automatic and tends to be **highly robust** to outliers and level shifts in the time series.


```{r X11, warning=F, message=F}

#Manufacture of electrical equipment: computer, electronic and optical products.
?elecequip

#TSA plot
autoplot(elecequip)

# seasonal effect fitting
elecequip %>% seas(x11="") -> fit


#A decomposition of the new orders index for electrical equipment.
autoplot(fit) +
  ggtitle("X11 decomposition of electrical equipment index")


#the trend-cycle component and the seasonally adjusted data, along with the original data.
autoplot(elecequip, series="Data") +
  autolayer(trendcycle(fit), series="Trend") +
  autolayer(seasadj(fit), series="Seasonally Adjusted") +
  xlab("Year") + ylab("New orders index") +
  ggtitle("Electrical equipment manufacturing (Euro area)") +
  scale_colour_manual(values=c("gray","blue","red"),
                      breaks=c("Data","Seasonally Adjusted","Trend"))
```

'Seasonally Adjusted' has less fluctuation than before.

Therefore 'Electrical equipment manufacturing' data is heavily influenced by the seasonal variation.

```{r X11_2, warning=F, message=F}
#Seasonal sub-series plot of the seasonal component from the X11 decomposition of the new orders index for electrical equipment.
fit %>% seasonal() %>% ggsubseriesplot() + ylab("Seasonal")
```

blue line : seasonal effect of each month.

black line : Time-series graphs with constants.

```{r X11_3, warning=F, message=F}
#fit %>% seasonal()
#fit %>% trendcycle()
#fit %>% remainder()
#fit %>% seasadj()
```

Given the output from the `seas()` function, 

`seasonal()` will extract the seasonal component, 

`trendcycle()` will extract the trend-cycle component, 

`remainder()` will extract the remainder component, and 

`seasadj()`will compute the seasonally adjusted time series.


# STL (Seasonal Decomposition of Time Series by Loess)

- STL is a versatile and robust method for decomposing time series.

- STL is an acronym for “Seasonal and Trend decomposition using Loess”, while **Loess is a method for estimating nonlinear relationships**. The STL method was developed by Cleveland, Cleveland, McRae, & Terpenning (1990).

- STL has several advantages over the classical, SEATS and X11 decomposition methods:

Unlike SEATS and X11, **STL will handle any type of seasonality, not only monthly and quarterly data.**

- The seasonal component is allowed to change **over time, and the rate of change can be controlled by the user.**

- **The smoothness of the trend-cycle can also be controlled by the user.**

- It can be **robust to outliers** (i.e., the user can specify a robust decomposition), so that occasional unusual observations will not affect the estimates of the trend-cycle and seasonal components. They will, however, affect the remainder component.

- On the other hand, STL has some disadvantages. In particular, it does **not handle trading day or calendar variation automatically**, and it **only provides facilities for additive decompositions.**

```{r air_STL}
air_stl=stl(AirPassengers,'periodic')
air_stl

plot(air_stl)
air_ss=seasadj(air_stl) #Returns seasonally adjusted data constructed by removing the seasonal component.

seasonal_index=sindexf(air_stl, h=24) # Returns vector containing the seasonal index for h future periods.
seasonal_index

autoplot(air_stl) # decomposition plot

forecast_air<-forecast::forecast(air_stl)

forecast_air

autoplot(forecast_air)
```

## Visualization of TSA, seasonal graphs using ggplot2
```{r air_STL_1}
data(a10)

?a10

head(a10)
str(a10)
glimpse(a10)

# ACF
ggAcf(a10)
ggAcf(a10, lag=48)

# PACF
ggPacf(a10)
ggPacf(a10, lag=48)


#Monthly sales of antidiabetic drugs in Australia.
autoplot(a10) +
  ggtitle("Antidiabetic drug sales") +
  ylab("$ million") +
  xlab("Year")

#Seasonal plot of monthly antidiabetic drug sales in Australia.
ggseasonplot(a10, year.labels=TRUE, year.labels.left=TRUE) +
  ylab("$ million") +
  ggtitle("Seasonal plot: antidiabetic drug sales")

#Polar seasonal plot of monthly antidiabetic drug sales in Australia.
ggseasonplot(a10, polar=TRUE) +
  ylab("$ million") +
  ggtitle("Polar seasonal plot: antidiabetic drug sales")



#Seasonal subseries plot of monthly antidiabetic drug sales in Australia.
ggsubseriesplot(a10) +
  ylab("$ million") +
  ggtitle("Seasonal subseries plot: antidiabetic drug sales")
```






