---
title: "time series2"
author: "hojae"
date: '2021 3 20 '
output:
  pdf_document: default
  html_document: default
---


## Moving average smoothing practice

Moving average at 3, 6, and 12 points for Airpassenger data
```{r MA}
data("AirPassengers")
m3=filter(AirPassengers, filter=rep(1/3,3))
m6=filter(AirPassengers, filter=rep(1/6,6))
m12=filter(AirPassengers, filter=rep(1/12,12))

plot(AirPassengers, main="AirPassengers: original data")
plot(m3, main="3-point moving average")
plot(m6, main="6-point moving average")
plot(m12, main="12-point moving average")
```

For data (dd1), try moving average smoothing and check the residual plot.

```{r dd1_MA4}
#install.packages("forecast")
library(forecast)

dd1=matrix(c(1342, 1442, 1252, 1343, 1425, 1362, 1456, 1272, 1243,
1359,1412, 1253, 1201, 1478, 1322, 1406, 1254, 1289, 1497, 1208))
dd1.ts=ts(data=dd1, start=c(2006,1), frequency=4)
dd1.ts

ff=filter(dd1.ts, filter=rep(1,4)/4, method="convolution", sides=1)
plot(dd1.ts, main="dd1.ts with simple moving average smoothing")
lines(ff, col="red", lty=2, lwd=2) # simple moving average smoothing
abline(h=mean(dd1), col="blue") # mean

res=ff[-1:-3,]-dd1.ts[-1:-3,] #remove first three values which are not able to be predicted.
tsdisplay(res, main="Residuals by MA(4) for dd1")
Box.test(res, type="Box-Pierce") # H0: TS is independent.
```

The residual shows no auto correlation after the model MA(4) is applied.


## double moving average

```{r dd1_double_MA4}
ff1=filter(dd1.ts, filter=rep(1,3)/3, method="convolution", sides=1)

ff2=filter(ff1, filter=rep(1,3)/3, method="convolution", sides=1)

plot(dd1.ts, main="dd1.ts with simple moving average smoothing")
lines(ff2, col="red", lty=2, lwd=2)
abline(h=mean(dd1), col="blue")

res=ff2[-1:-3,]-dd1.ts[-1:-3,]
tsdisplay(res, main="Residuals by double MA smoothing")
Box.test(res, type="Box-Pierce") # H0: TS is independent.
```
The residual shows no auto correlation after the double moving average model is applied.


## double moving average (2nd method)

```{r dd1_double_MA4_2}
mm1=ma(dd1.ts, order=4, centre=F) #MA(4) without a center
mm3=ma(dd1.ts, order=3, centre=T) #MA(3) with a center
mm4=ma(mm3, order=3, centre=T) #double MA(3) with a center

res3=mm4-dd1.ts

Box.test(res3)

plot(forecast(mm1, h=2),main="forecast simple MA") # forecast
plot(forecast(mm4, h=2),main="forecast simple MA") # forecast

f1=forecast(mm1, h=1)
f1

f4=forecast(mm4, h=1)
f4

accuracy(f1)
accuracy(f4)
```


## Simple exponential smoothing practice

Let’s fit the simple exponential smoothing practice with alpha=0.1

```{r expo_smoothing}
ho=HoltWinters(dd1.ts, alpha=0.1, beta=F, gamma=F) #exponential smoothing # (beta=F, gamma=F): no trend and no seasonal effect
ho
ho$fitted #fitting value
plot(ho)
```


Let’s fit the simple exponential smoothing practice with the optimized alpha

```{r expo_smoothing_alpha}
ha=HoltWinters(dd1.ts,beta=F, gamma=F) #exponential smoothing
ha
ha$fitted #fitting value
plot(ha) # the red line is the fitted value
```

Try a residual analysis:

```{r residual_smoothing}
fa=forecast(ha)
names(fa)
fa$fitted
fa$residuals
tsdisplay(fa$residual)
Box.test(fa$residual, type="Box-Pierce") # p-value = 0.1496 : independent
hist(fa$residuals, 10, probability=T, col="light blue", xlab="Time", ylab="Residuals", main="Histogram of Residuals")
points(density (fa$residuals[-1], bw=30), type='l', col="red", lwd=2)
title("\n \n - simple exponential smoothing")
```

The forecasting result:

```{r forecast_smoothing}
fa=forecast(ha, h=4) # forecasting four time points
fa
accuracy(fa)

plot(forecast(ha, h=4)) # shows 80%, 95% significant level for forecasting
lines(fa$fitted, col="red", lty=2, lwd=2)

plot(forecast(ha, h=4, fan=T)) # shows 51-99% significant level for forecasting
lines(fa$fitted, col="red", lty=2, lwd=2)
```






